The Changing Face of Pediatric Pulmonary Exacerbations in CF (NACFC 2023)

### 2023-03-29 Scott Sagel/PEx in CF/nacfc_2023.qmd
Update nacfc_2023.qmd
tidyverse styling

chco = read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CHCO.xlsx",na = c("","NULL"))
cnh = read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CNH.xlsx")

@@

chco$`P. aeruginosa` = coalesce(chco$`P. aeruginosa mucoid`,chco$`P. aeruginosa non-mucoid`)
chco = chco %>% rename(
  Race = `Race (White/Black/Other)`,Ethnicity = `Hispanic/Latinx`

@@

  mutate(Genotypes1 = case_when(

@@

  `FEV1 % Predicted at Admit` = suppressWarnings(as.numeric(`FEV1 % Predicted at Admit`))) %>%
  unite(Genotype,Genotypes1,Genotypes2,sep = "/")

@@

cnh = cnh %>%
  rename(`# Days in Admit`=`Total hospitalization days`,
         `Age at Admit`=`Age at admission`,
         `FEV1 % Predicted at Discharge`=`End of treamtent ppFEV1`,
         `Unique number` = `Study ID`) %>%
  mutate(Site = "CNH",

@@

           None =="Checked" ~"No"

@@

           MRSA =="Checked" ~ TRUE

@@

           MSSA =="Checked" ~ TRUE

@@

           `P. aeruginosa` =="Checked" ~ TRUE

@@

           `CFTR genotype, groups` =="F508del heterozygous" ~ "F508del/Other",
           `CFTR genotype, groups` =="F508del homozygous"~"F508del/F508del",
           `CFTR genotype, groups` =="Other"~"Other/Other"

@@

           `Tezacaftor/ivacaftor (Symdeko)`=="Checked" ~ "Other",
           `Elexacaftor/tezacaftor/ivacaftor (Trikafta)`=="Checked" ~ "Trikafta",

@@

         `FEV1 % Predicted at Admit` = as.numeric(`PEx ppFEV1`))

@@

df = full_join(chco,cnh)

@@

df$`Modulator at Time of Admit` = factor(df$`Modulator at Time of Admit`,
                                         levels = c("No","Yes"),ordered = T)
df$`Specific Modulator at Admit`[df$`Modulator at Time of Admit`=="No"] = "None"
df$`Specific Modulator at Admit` = factor(tolower(df$`Specific Modulator at Admit`),
                                          levels = c("kalydeco","none","orkambi","symdeko","trikafta"),
                                          labels = c("Other","None","Other","Other","Trikafta"))
df$Race = factor(df$Race,levels = c("Amer. Indian or Alaskan Naitive","Asian",
                                    "Black","Black or African American","Black/ AA",
                                    "Decline to Answer","Hispanic","Other",
                                    "Unknown / Not Reported","White"),
                 labels = c("Other","Other","Black","Black","Black",
                            "Unknown / Not Reported","Other","Other",
                            "Unknown / Not Reported","White"))
df$Ethnicity = factor(df$Ethnicity,levels = c("FALSE","Hispanic or Latino","HLS",
                                              "NOT Hispanic or Latino","NOT HLS",
                                              "TRUE","UNKNOWN"),
                      labels = c("NOT Hispanic or Latino","Hispanic or Latino",
                                 "Hispanic or Latino","NOT Hispanic or Latino",
                                 "NOT Hispanic or Latino","Hispanic or Latino",
                                 "Unknown"))
bugs = c("P. aeruginosa","P. aeruginosa mucoid","P. aeruginosa non-mucoid","Staph aureus","MRSA")
df[,bugs] = lapply(df[,bugs],function(c){
  c[is.na(c)] = "No"
  c = factor(c,levels = c("No",TRUE),labels = c("No","Yes"),ordered = T)

@@

@@

summ = df %>% group_by(`Unique number`,Site,Year) %>%

@@

  summarise(`# of Hospitalizations per Person` = n(),
            across(c(Sex:Genotype,`Specific Modulator at Admit`),~first(na.omit(.x))),
            across(c(`Modulator at Time of Admit`,all_of(bugs)),~max(.x,na.rm = T)),
            across(all_of(c("Age at Admit","# Days in Admit","Change in FEV1",
                            "FEV1 % Predicted at Admit","FEV1 % Predicted at Discharge")),
                   ~mean(.x,na.rm=T)),
            .groups = "drop")  %>%

@@

                           levels=c("F508del/F508del","F508del/Other",
                                    "Other/F508del","Other/Other"),
                           labels = c("F508del/F508del","F508del/Other",
                                      "F508del/Other","Other/Other")

@@

  select(`Unique number`, `Site`,Year,`# of Hospitalizations per Person`,`# Days in Admit`,`Age at Admit`,
         `Change in FEV1`,`FEV1 % Predicted at Admit`,`FEV1 % Predicted at Discharge`,
         `Modulator at Time of Admit`,`Specific Modulator at Admit`,
         all_of(bugs),Sex:Genotype)

@@

```{r}
#| label: tbl-all
#| tbl-cap: Summary by Year
#| message: false
summ %>%
  select(-`Unique number`,-Site) %>%
  tbl_summary(by=Year,
              type = list(`# of Hospitalizations per Person` ~ 'continuous'),
              statistic = list(
                c("# of Hospitalizations per Person","# Days in Admit") ~ "{median} ({p25}, {p75})",
                c("Age at Admit","Change in FEV1","FEV1 % Predicted at Admit",
                  "FEV1 % Predicted at Discharge") ~ "{mean} ({sd})"
              ),
              missing_text = "(Missing)") %>%
  add_ci(pattern = "{stat} [{ci}]") %>%
  add_p(test = list(
    `Age at Admit`~"t.test",
    `Change in FEV1`~"t.test"
  )) %>%
  separate_p_footnotes() %>%
  as_gt() %>%
  as_raw_html()
```

```{r}
#| label: tbl-site
#| tbl-cap: Stratified by Site
#| message: false
summ %>%
  select(-`Unique number`) %>%
  tbl_strata(
    strata = Site,
    .tbl_fun =
      ~ .x %>%
      tbl_summary(by=Year,
              type = list(`# of Hospitalizations per Person` ~ 'continuous'),
              statistic = list(
                c("# of Hospitalizations per Person","# Days in Admit") ~ "{median} ({p25}, {p75})",
                c("Age at Admit","Change in FEV1","FEV1 % Predicted at Admit",
                  "FEV1 % Predicted at Discharge") ~ "{mean} ({sd})"
              ),
              missing_text = "(Missing)") %>%
  add_ci(pattern = "{stat} [{ci}]") %>%
  add_p(test = list(
    `Age at Admit`~"t.test",
    `Change in FEV1`~"t.test"
  )) %>%
  separate_p_footnotes()
  ) %>%
  as_gt() %>%
  as_raw_html()
```

>>>>

chco <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CHCO.xlsx", na = c("", "NULL"))
cnh <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CNH.xlsx")

@@

chco$`P. aeruginosa` <- coalesce(chco$`P. aeruginosa mucoid`, chco$`P. aeruginosa non-mucoid`)
chco <- chco %>%
  rename(
    Race = `Race (White/Black/Other)`, Ethnicity = `Hispanic/Latinx`

@@

  mutate(
    Genotypes1 = case_when(

@@

    `FEV1 % Predicted at Admit` = suppressWarnings(as.numeric(`FEV1 % Predicted at Admit`))
  unite(Genotype, Genotypes1, Genotypes2, sep = "/")

@@

cnh <- cnh %>%
  rename(
    `# Days in Admit` = `Total hospitalization days`,
    `Age at Admit` = `Age at admission`,
    `FEV1 % Predicted at Discharge` = `End of treamtent ppFEV1`,
    `Unique number` = `Study ID`
  mutate(
    Site = "CNH",

@@

      None == "Checked" ~ "No"

@@

      MRSA == "Checked" ~ TRUE

@@

      MSSA == "Checked" ~ TRUE

@@

      `P. aeruginosa` == "Checked" ~ TRUE

@@

      `CFTR genotype, groups` == "F508del heterozygous" ~ "F508del/Other",
      `CFTR genotype, groups` == "F508del homozygous" ~ "F508del/F508del",
      `CFTR genotype, groups` == "Other" ~ "Other/Other"

@@

      `Tezacaftor/ivacaftor (Symdeko)` == "Checked" ~ "Other",
      `Elexacaftor/tezacaftor/ivacaftor (Trikafta)` == "Checked" ~ "Trikafta",

@@

    `FEV1 % Predicted at Admit` = as.numeric(`PEx ppFEV1`)
  )

@@

df <- full_join(chco, cnh)

@@

df$`Modulator at Time of Admit` <- factor(df$`Modulator at Time of Admit`,
  levels = c("No", "Yes"), ordered = T
)
df$`Specific Modulator at Admit`[df$`Modulator at Time of Admit` == "No"] <- "None"
df$`Specific Modulator at Admit` <- factor(tolower(df$`Specific Modulator at Admit`),
  levels = c("kalydeco", "none", "orkambi", "symdeko", "trikafta"),
  labels = c("Other", "None", "Other", "Other", "Trikafta")
)
df$Race <- factor(df$Race,
  levels = c(
    "Amer. Indian or Alaskan Naitive", "Asian",
    "Black", "Black or African American", "Black/ AA",
    "Decline to Answer", "Hispanic", "Other",
    "Unknown / Not Reported", "White"
  labels = c(
    "Other", "Other", "Black", "Black", "Black",
    "Unknown / Not Reported", "Other", "Other",
    "Unknown / Not Reported", "White"
  )
)
df$Ethnicity <- factor(df$Ethnicity,
  levels = c(
    "FALSE", "Hispanic or Latino", "HLS",
    "NOT Hispanic or Latino", "NOT HLS",
    "TRUE", "UNKNOWN"
  labels = c(
    "NOT Hispanic or Latino", "Hispanic or Latino",
    "Hispanic or Latino", "NOT Hispanic or Latino",
    "NOT Hispanic or Latino", "Hispanic or Latino",
    "Unknown"
  )
)
bugs <- c("P. aeruginosa", "P. aeruginosa mucoid", "P. aeruginosa non-mucoid", "Staph aureus", "MRSA", "S. maltophilia")
df[, bugs] <- lapply(df[, bugs], function(c) {
  c[is.na(c)] <- "No"
  c <- factor(c, levels = c("No", TRUE), labels = c("No", "Yes"), ordered = T)

@@

df$`S. maltophilia`[df$Site == "CNH"] <- NA

@@

summ <- df %>%
  group_by(`Unique number`, Site, Year) %>%

@@

  summarise(
    `# of Hospitalizations per Person` = n(),
    across(c(Sex:Genotype, `Specific Modulator at Admit`), ~ first(na.omit(.x))),
    across(c(`Modulator at Time of Admit`, all_of(bugs)), ~ max(.x, na.rm = T)),
    across(
      all_of(c(
        "Age at Admit", "# Days in Admit", "Change in FEV1",
        "FEV1 % Predicted at Admit", "FEV1 % Predicted at Discharge"
      )),
      ~ mean(.x, na.rm = T)
    .groups = "drop"

@@

    levels = c(
      "F508del/F508del", "F508del/Other",
      "Other/F508del", "Other/Other"
    labels = c(
      "F508del/F508del", "F508del/Other",
      "F508del/Other", "Other/Other"
    )

@@

  select(
    `Unique number`, `Site`, Year, `# of Hospitalizations per Person`, `# Days in Admit`, `Age at Admit`,
    `Change in FEV1`, `FEV1 % Predicted at Admit`, `FEV1 % Predicted at Discharge`,
    `Modulator at Time of Admit`, `Specific Modulator at Admit`,
    all_of(bugs), Sex:Genotype
  )

@@

```{r}
#| label: tbl-all
#| tbl-cap: Summary by Year
#| message: false
summ %>%
  select(-`Unique number`, -Site) %>%
  tbl_summary(
    by = Year,
    type = list(`# of Hospitalizations per Person` ~ "continuous"),
    statistic = list(
      c("# of Hospitalizations per Person", "# Days in Admit") ~ "{median} ({p25}, {p75})",
      c(
        "Age at Admit", "Change in FEV1", "FEV1 % Predicted at Admit",
        "FEV1 % Predicted at Discharge"
      ) ~ "{mean} ({sd})"
    ),
    missing_text = "(Missing)"
  ) %>%
  add_ci(pattern = "{stat} [{ci}]") %>%
  add_p(test = list(
    `Age at Admit` ~ "t.test",
    `Change in FEV1` ~ "t.test"
  )) %>%
  separate_p_footnotes() %>%
  as_gt() %>%
  as_raw_html()
```

```{r}
#| label: tbl-site
#| tbl-cap: Stratified by Site
#| message: false
summ %>%
  select(-`Unique number`, -`S. maltophilia`) %>%
  tbl_strata(
    strata = Site,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = Year,
          type = list(`# of Hospitalizations per Person` ~ "continuous"),
          statistic = list(
            c("# of Hospitalizations per Person", "# Days in Admit") ~ "{median} ({p25}, {p75})",
            c(
              "Age at Admit", "Change in FEV1", "FEV1 % Predicted at Admit",
              "FEV1 % Predicted at Discharge"
            ) ~ "{mean} ({sd})"
          ),
          missing_text = "(Missing)"
        ) %>%
        add_ci(pattern = "{stat} [{ci}]") %>%
        add_p(test = list(
          `Age at Admit` ~ "t.test",
          `Change in FEV1` ~ "t.test"
        )) %>%
        separate_p_footnotes()
  ) %>%
  as_gt() %>%
  as_raw_html()
```
